Nonlinear Dendritic Coincidence Detection for Supervised Learning

### 2021-03-30 notes/draft_article.tex
.#

		\subsection{Supervised Learning in a Linear Classification Task}

		Next, we investigated if the observed differences would also improve
		the performance in an actual supervised learning task.
		For this purpose, we constructed presynaptic basal input as illustrated
		in Fig.~\ref{fig:illustration_classification}. Written in vector form,
		each sample from the basal input was generated from the following expression:
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			\mathbf{x}_p(t) = \mathbf{b} + \mathbf{a}\left(c(t) + \sigma_a \zeta_a(t) \right) 
			+ s \cdot \sum_{i=1}^{N_{dist}} \zeta_{dist,i}(t) \mathbf{v}_{dist,i} \; .
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		Here, $\mathbf{b}$ is a random vector drawn uniformly from
		$(0,1)^N$, $\mathbf{a}$ is random unit vector as introduced in 
		Section~\ref{sect:alignment}, $c(t)$ is a binary variable drawn 
		from $\{-0.5,0.5\}$ with equal probability and $\zeta_a(t)$ and the
		$\zeta_{dist,i}(t)$ are independent random Gaussian variables.
		Finally, the set of $\mathbf{v}_{dist,i}$ forms a randomly generated
		orthogonal basis of $N_{dist}$ unit vectors which---just as in 
		Section~\ref{sect:alignment}---are also orthogonal to $\mathbf{a}$.
		As indicated by the time dependence, the Gaussian and binary random
		variables were drawn for each time step. The vectors
		$\mathbf{b}$, $\mathbf{a}$, and $\mathbf{v}_{dist,i}$ were generated
		once before the beginning of a simulation run.

		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{figure}
			\centering
			\includegraphics[width=0.75\columnwidth]{illustration_classification}
			\caption{{\bf Input Space for a Simple Linear Classification Task.}
			Two clusters of presynaptic basal activity were generated from 
			multivariate Gaussian distributions. Here, $s$ denotes the standard
			deviation orthogonal to the normal vector of the classification
			hyperplane.}
			\label{fig:illustration_classification}
		\end{figure}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

		\subsection{Supervised Learning in a Linear Classification Task}
		\label{sect:classification}

		Next, we investigated if the observed differences would also improve
		the performance in an actual supervised learning task.
		For this purpose, we constructed presynaptic basal input as illustrated
		in Fig.~\ref{fig:illustration_classification}.
		Written in vector form,
		each sample from the basal input was generated from the following expression:
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{equation}
			\mathbf{x}_p(t) = \mathbf{b} + \mathbf{a}\left(c(t) + \sigma_a \zeta_a(t) \right) 
			+ s \cdot \sum_{i=1}^{N_{dist}} \zeta_{dist,i}(t) \mathbf{v}_{dist,i} \; .
		\end{equation}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		Here, $\mathbf{b}$ is a random vector drawn uniformly from
		$(0,1)^N$, $\mathbf{a}$ is random unit vector as introduced in 
		Section~\ref{sect:alignment}, $c(t)$ is a binary variable drawn 
		from $\{-0.5,0.5\}$ with equal probability and $\zeta_a(t)$ and the
		$\zeta_{dist,i}(t)$ are independent random Gaussian variables with 
		zero mean and unit variance. 
		Hence, $\sigma_a$ simply denotes the standard deviation of each Gaussian
		cluster along the direction of the normal vector $\mathbf{a}$ and was
		set to $\sigma_a = 0.25$. 
		Finally, the set of $\mathbf{v}_{dist,i}$ forms a randomly generated
		orthogonal basis of $N_{dist}$ unit vectors which---just as in 
		Section~\ref{sect:alignment}---are also orthogonal to $\mathbf{a}$.
		The free parameter $s$ parameterized
		the standard deviation along this subspace orthogonal to $\mathbf{a}$.
		As indicated by the time dependence, the Gaussian and binary random
		variables were drawn for each time step. The vectors
		$\mathbf{b}$, $\mathbf{a}$, and $\mathbf{v}_{dist,i}$ were generated
		once before the beginning of a simulation run.
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{figure}
			\centering
			\includegraphics[width=0.75\columnwidth]{illustration_classification}
			\caption{{\bf Input Space for a Simple Linear Classification Task.}
				Two clusters of presynaptic basal activity were generated from 
				multivariate Gaussian distributions. Here, $s$ denotes the standard
				deviation orthogonal to the normal vector of the classification
				hyperplane.}
			\label{fig:illustration_classification}
		\end{figure}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

		For the classification task, we created two output neurons receiving
		the same basal presynaptic input, while the top-down input encoded
		the correct linear classification in a one-hot scheme, that is
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{align}
			x_{d,0}(t) &= 1 - \Theta\left( \left(\mathbf{x}_p(t) - \mathbf{b}\right)^T \mathbf{a}\right) \\
			x_{d,1}(t) &= \Theta\left( \left(\mathbf{x}_p(t) - \mathbf{b}\right)^T \mathbf{a}\right) \; ,
		\end{align}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		where $\Theta(x)$ is the Heaviside step function.

		As in the previous experiment, we ran a full simulation until 
		all dynamic variables reached a stationary state. After this,
		a test run without plasticity and with the apical input turned off 
		was used to evaluate the classification	performance. For each sample,
		the index of the neuron with the highest activity was used
		as the predicted class. Accuracy was then calculated as the fraction
		of correctly classified samples.
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{figure}
			\includegraphics[width=\columnwidth]{classification_dimension_scaling}
			\caption{{\bf Binary Classification Accuracy.}
			Fraction of correctly classified patterns as illustrated in
			Fig.~\ref{fig:illustration_classification}, see 
			Section~\ref{sect:classification}.¸}
		\label{fig:classification_accuracy}
		\end{figure}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

		The resulting accuracy as a function of $N_{dist}$ and $s$
		is shown in Fig.~\ref{fig:classification_accuracy}.
		Albeit differences were small, the compartment model did
		show a better overall accuracy for the tested parameter range.

		\subsection{Non-Hebbian Learning Rules}
		\label{sect:non-hebbian}

		To be continued, see Fig.~\ref{fig:classification_accuracy_bcm}.

		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
		\begin{figure}
			\includegraphics[width=\columnwidth]{classification_dimension_scaling_bcm}
			\caption{{\bf Binary Classification Accuracy, BCM Rule.}
				Fraction of correctly classified patterns as illustrated in
				Fig.~\ref{fig:illustration_classification}, after training with
				a BCM-like learning rule.¸}
			\label{fig:classification_accuracy_bcm}
		\end{figure}
		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

		\section{Discussion}
